---
title : 'Krysten Thompson w271: Homework 6'
author: "Professor Jeffrey Yau"
output: 
  pdf_document:
  toc: true
  number_sections: true
fontsize: 11pt
geometry: margin=1in
---

In this homework, you are asked to conduct Time Series EDA, develop a Time Trend Model, conduct model diagnostic analysis, and use the model to make forecasts.

Load the file **"bls_unemployment.csv"**
This file contains the monthly unemployment rate in the United States from January 2007 to January 2017. *Note that these data are NOT seasonally adjusted.*

1. Load the csv file into a data.frame, calling it *df*, and examine the structure of the "raw" series after you load it into a data.frame.


```{r}
library(astsa)
library(ggplot2)
library(knitr)
library(timeSeries)

opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

```{r}
df <- read.csv('bls_unemployment.csv')
head(df)
```

```{r}
str(df)
```

2. Convert it into a R time-series object, and examine the structure of the series after you convert the *df* into a time series object. What is the difference between the two structures?

```{r}
class(df)
```

```{r}
df <- ts(df, start = c(2007, 1), freq = 12)
class(df)
```

```{r}
str(df)
```

```{r}
head(df)
```

```{r}
#describe(df) 

#because I changed 'df' to ts, the describe function won't run
```

Suppose we want to approximate the unemployment rate between 2010 and 2017 using a linear time trend model. For this exercise, feel free to modify this dataset but be sure that you explain what you did and why.

**I created a subset of the dataset, specifically data for 2010 through 2017. I created a new variable 'df.2010_2017'.**

**First, I plotted the variable with no modification. But this showed the in-year trend (e.g. monthly data).**

**Then I plotted the data and called 'aggregate' so that the plot would reflect a smooth linear time trend line.**

```{r}
df.2010_2017 <- window(df[, 4], start = c(2010, 1), end = c(2017, 1))
plot(df.2010_2017, ylab = 'Unemployment Rate by Year', xlab = 'Year')
```

```{r}
plot(aggregate(df.2010_2017))
```


3. Conduct EDA on the series.

```{r}
start(df); end(df); frequency(df)
```

```{r}
plot(df[,4], ylab = 'Unemployment %', xlab = 'Year')
```

```{r}
boxplot(df[,4] ~ cycle(df), ylab = '% Unemployment', xlab = 'Month', main = 'Boxplots of Unemployment by Month Aggregated 2007 - 2017')
```

```{r}
sum(is.na(df))
```


4. Create a linear time trend model by regressing the unemployment rate on time. Interpret the model results. 
  
```{r}
mod.lin <- lm(df[,4] ~ Year, data=df)
summary(mod.lin)
```

**Coefficient for Year $(-0.15)$ suggests that for every year after 2007, unemployment decreases by ~$15\%$ per year.**

**The intercept value is high likely due to the Great Recession starting in late 2007 and into 2008. The significant Standard Error of $115$ is a result of 2007 being low and then the unemployment rate spiking in 2008 and extending into 2010.**

**While the p-value of $0.0085$ is <$0.05\%$, it's important to note the Multiple $R^2$ value of $0.0568$. Approximately $6\%$ of the variance in unemployment rate is due to time. This is because there are many confounding variables not included in this model that better explain the unemployment rate.**
  
5. Examine the residuals of this model as you would if it were a classical linear model. In addition, generate ACF and PACF plots of the residuals.

**Because I discussed the residuals of summary(model) above, I created a residuals plot below using decompose and plotting "random".**

```{r}
df.decom <- decompose(df, "multiplicative")
plot(ts(df.decom$random[7:115]), main= "Residuals Plot")
```

```{r}
acf(df.decom$random[7:115], main="ACF Plot for Residuals")
```

```{r}
pacf(df.decom$random[7:115], main = "PACF Plot for Residuals")
```

**I wanted to play around and see what happened when I plotted "Trends" and "Seasonal".**

```{r}
plot(ts(df.decom$trend[,4]), main= "Trends Plot")
```

```{r}
plot(ts(df.decom$seasonal[7:115]), main= "Seasonal Plot")

```


6. Use this model to predict the unemployment rate in 2018 January (i.e. 12 months from the end of the sample). Do the result make sene? How about a prediction of the unemployment rate in 2020 December?

**Gerard, I couldn't figure this out after extensive online searching. Gave it a try using 'auto.arima' but couldn't get it to print for Jan 2018.**

```{r}
# 
library(forecast)
# 
# frcst_Jan <- forecast(mod.lin, n.ahead=12)
# frcst_Jan
```

```{r}
value <- data.frame(df[, 4])
.months <- ts(value, frequency=12)
fit <- auto.arima(.months)
pred <- predict(fit, n.ahead=14)
pred
```








