---
title: 'W271 Homework 9'
author: "Carlos Sancini"
geometry: margin=1in
output:
  pdf_document: null
  number_sections: yes
  html_document: default
  toc: yes
fontsize: 11pt
---

# Develop a Vector Autoregressive Model:

Use series `series01_liveSession_wk10.csv` and build a VAR model. You will have to examine the data, conduct EDA, use `VARselect` to choose a model that minimize `SC` (which is also `BIC`), estimate the chosen model, conduct residual diagnostic, test model assumptions, and make a 3-step ahead forecast.

Setup code:

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE,warning=FALSE, message=FALSE)
```

Load required libraries:

```{r, message=FALSE, warning=FALSE}

# Clean up the workspace before we begin
rm(list = ls())

# Load required libraries
library(car)
library(dplyr)
library(Hmisc)
library(gridExtra)
library(forecast)
library(fpp2)
library(astsa)
library(xts)
library(vars)

# Set working directory
wd = "D:\\GDrive\\W271\\repo\\hw\\hw09"
setwd(wd)

```


Load dataset, convert to time series and analyze structure:

```{r}
df = read.csv("series01_liveSession_wk10.csv", encoding = "UTF-8", header = TRUE, sep=",")

# Examine the data structure
rbind(head(df, 5), some(df, 5), tail(df, 5))
str(df)
describe(df)

# convert to time series
xts_idx = seq(as.Date("2005/1/1"), by = "month", length.out = nrow(df))
ts = ts(df, start = c(2005,1), frequency = 12)
head(ts)
tail(ts)

autoplot(ts, facets=TRUE) + xlab("Year") + ggtitle("Series 1 & 2 Plots")


```

EDA:

Autocorrelation analisys

```{r}

# check autocorrelation
a1 = ggAcf(ts[,1], lag=24) +
  ggtitle("Series 1")

# check partial autocorrelation
a2 = ggPacf(ts[,1], lag=24) +
  ggtitle("Series 1")

# check autocorrelation
a3 = ggAcf(ts[,2], lag=24) +
  ggtitle("Series 2")

# check partial autocorrelation
a4 = ggPacf(ts[,2], lag=24) +
  ggtitle("Series 2")

grid.arrange(a1, a2, a3, a4, ncol=2)

```

Seasonality analisys

```{r}
# check for seasonality
sb1 = ggplot() + 
  ggtitle("Seasonal boxplot: Series 1 & 2") +
  geom_boxplot(mapping = aes(x = factor(cycle(ts[,1])), y = ts[,1]), color="black", fill="gray") +
  labs(x = "", y =  "") +
  scale_y_continuous(expand = c(0,0))
sb2 = ggplot() +
  geom_boxplot(mapping = aes(x = factor(cycle(ts[,2])), y = ts[,2]), color="black", fill="gray") +
  labs(x = "Month", y =  "") +
  scale_y_continuous(expand = c(0,0))
grid.arrange(sb1, sb2, ncol=1)
```

Contemporaneous correlation analisys

```{r}
scatterplotMatrix(~ ts[,1] + ts[,2])
title("Contemporaneous Correlation of Series 1 & 2")
```

```{r}

cat("Correlation Matrix: ", cor(ts[,1],ts[,2]))

ccf(ts[,1],ts[,2])

```


MODEL SELECTION:

```{r}
VARselect(ts, lag.max = 8, type = "both")
```

```{r}
var.fit = VAR(ts, p = 1, type = "both")
summary(var.fit)
```

```{r}
roots(var.fit)
```

DIAGNOSTIC TESTING: 

```{r}
# Test of normality: null hypothesis = normally distributed
normality.test(var.fit, multivariate.only = TRUE)

# Test of no serial correlation: null hypothesis = no serial correlation
serial.test(var.fit, lags.pt = 12, type = "PT.asymptotic")
#plot(var.fit.ptasy)

# Test of the absence of ARCH effect: null hypothesis = homoscedastic
arch.test(var.fit)
```

FORECAST: 

```{r}
forecast(var.fit) %>% autoplot() + 
  ggtitle("VAR model forecast") +
  labs(x = "Year", y =  "") 
```