---
title: 'Krysten Thompson - w271: Homework 9 (Due: Week 11)'
author: "Professor Jeffrey Yau"
geometry: margin=1in
output:
  pdf_document: null
  number_sections: yes
  html_document: default
  toc: yes
fontsize: 11pt
---

# Develop a Vector Autoregressive Model:

Use series `series01_liveSession_wk10.csv` and build a VAR model. You will have to examine the data, conduct EDA, use `VARselect` to choose a model that minimize `SC` (which is also `BIC`), estimate the chosen model, conduct residual diagnostic, test model assumptions, and make a 3-step ahead forecast.

```{r, warning=FALSE, message=FALSE}

library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

```{r warning=FALSE, message=FALSE}
library(car)
library(dplyr)
library(Hmisc)
library(gridExtra)
library(forecast)
library(fpp2)
library(astsa)
library(xts)
library(vars)
```

```{r}
d <- read.csv("series01_liveSession_wk10.csv", header = TRUE, sep=",")
```

```{r}
str(d)
```

```{r}
head(d,3)
tail(d,3)
```

```{r}
describe(d)
```

```{r}
values = seq(from = as.Date("2000-01-01"), to = as.Date("2009-12-31"), by = 'month')
```

```{r}
d$date <- values
```

```{r}
head(d)
```

```{r}
d <- d[,c(3,1,2)]
head(d,3)
```

```{r}
d.ts <- ts(d)
head(d.ts)
```

```{r}
#Plot series data
plot.ts(d.ts, main="Plot of Series 1 and Series 2")
```

```{r}
idx <- seq(from = as.Date("2019-01-01"), to = as.Date("2019-04-30"), by = 'day')
```


```{r}

# check autocorrelation and partial autocorrelation for each series
s1 <- ggAcf(d.ts[,2], lag=24) + ggtitle("ACF Series 1")

s1p <- ggPacf(d.ts[,2], lag=24) + ggtitle("PACF Series 1")

s2 <- ggAcf(d.ts[,3], lag=24) + ggtitle("ACF Series 2")

s2p <- ggPacf(d.ts[,3], lag=24) + ggtitle("PACF Series 2")

grid.arrange(s1, s1p, s2, s2p, ncol=2)
```

```{r}
scatterplotMatrix(~ d.ts[,2] + d.ts[,3])
title("Contemporaneous Correlation of Series 1 & 2")
```

Select optimal number of lags:

```{r}
#I tested many lag.max (all the way up to 24) and it kept coming back to 1
VARselect(d.ts, lag.max = 5, type = "both")
```


**In a next step, the VAR model is estimated with the function VAR() and as deterministic regressors a constant is included.**

```{r}
var.fit <- VAR(d.ts, p = 1, type = "both")
summary(var.fit)
```

**The moduli of the eigenvalues of the companion matrix are all less than one.**

```{r}
roots(var.fit)
```

# Diagnostic Testing: 

```{r}
# Test of normality:
normality.test(var.fit, multivariate.only = TRUE)

# Test of no serial correlation:
serial.test(var.fit, lags.pt = 12, type = "PT.asymptotic")
#plot(var.fit.ptasy)

# Test of the absence of ARCH effect:
arch.test(var.fit)
```

# Forecast:

```{r}
forecast(var.fit) %>% autoplot() +
  ggtitle("Var Model Forecast") + xlab("Year")  + ylab("Value")

```








